How do lifestyle and poverty influence BMI across different age groups and genders

### 2024-12-03 paper/paper.qmd
some updates. still need: paper and datasheet

The dataset used in this analysis is derived from the US National Health and Nutrition Examination Survey (NHANES), version 2.1.0, published in July 2015. NHANES is a long-running study conducted by the US National Center for Health Statistics (NCHS) that has been gathering health and nutrition data since the early 1960s. Since 1999, approximately 5,000 individuals from various age groups have been interviewed annually in their homes and undergone health examinations at mobile examination centers (MEC). The dataset contains 10,000 observations and 76 variables. The data used here was originally compiled by Michelle Dalrymple from Cashmere High School and Chris Wild from the University of Auckland for educational purposes.

For the current study, the data was cleaned to focus on variables pertinent to the analysis of BMI. Specifically, variables such as BMI, poverty index, physical activity days, sleep hours, gender, and age were retained. After cleaning the missing values in the dataset, 3,573 observations remained. The dataset was prepared, cleaned, and analyzed using R (R Core Team, 2022) with the following libraries: opendatatoronto (Gelfand, 2022) for accessing the data, tidyverse (Wickham et al., 2019), dplyr (Wickham et al., 2023) for data manipulation, and ggplot2 (Wickham, 2016) for visualizations. Additionally, knitr (Xie, 2023a) was used for report generation, and styler (Müller et al., 2024) ensured the R code was properly styled.

@@

# Create the summary table using kable
summary_table %>%
  kable(caption = "Summary statistics for variables in the NHANES dataset.")
```

```{r}
#| label: fig-bmi-logbmi
#| fig-cap: "Distributions of BMI and log(BMI) in the NHANES dataset."
#| echo: false
#| fig.width: 8
#| fig.height: 4

@@

# Figure 1: BMI and log(BMI)
figure_bmi <- (hist_bmi | hist_log_bmi) +
  plot_annotation(caption = "Figure 1: Distributions of BMI and log(BMI) in the NHANES dataset.",
                  theme = theme(plot.caption = element_text(hjust = 0.5, size = 12)))
figure_bmi

@@

```{r}
#| label: fig-predictor-distributions
#| fig-cap: "Distributions of predictor variables in the NHANES dataset, including Poverty, Physical Activity Days, Age, Sleep Hours, and Gender."
#| echo: false
#| fig.width: 8
#| fig.height: 12

@@

```{r}
#| echo: false
#| label: tbl-modelresults
#| tbl-cap: "Summary statistics for variables in the NHANES dataset."

@@

```{r}
#| label: fig-predictor-bmi
#| fig-cap: "Scatter plots of predictors vs. log(BMI) with fitted regression lines."
#| fig.width: 8
#| fig.height: 12

# Scatter plot for Age vs. log(BMI)
plot_age_bmi <- ggplot(nhanes_data, aes(x = Age, y = log(BMI))) +
  geom_point(size = 0.3) +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(title = "Age vs. log(BMI)", x = "Age", y = "log(BMI)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12), axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# Scatter plot for Gender vs. log(BMI)
plot_gender_bmi <- ggplot(nhanes_data, aes(x = Gender, y = log(BMI))) +
  geom_point() +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(title = "Figure17: Gender vs. log(BMI)", x = "Gender", y = "log(BMI)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14), axis.title = element_text(size = 12), axis.text = element_text(size = 10))

# Scatter plot for Poverty vs. log(BMI)
plot_poverty_bmi <- ggplot(nhanes_data, aes(x = Poverty, y = log(BMI))) +
  geom_point() +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(title = "Figure18: Poverty vs. log(BMI)", x = "Poverty", y = "log(BMI)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14), axis.title = element_text(size = 12), axis.text = element_text(size = 10))

# Scatter plot for SleepHrsNight vs. log(BMI)
plot_sleep_bmi <- ggplot(nhanes_data, aes(x = SleepHrsNight, y = log(BMI))) +
  geom_point() +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(title = "Figure19: SleepHrsNight vs. log(BMI)", x = "SleepHrsNight", y = "log(BMI)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14), axis.title = element_text(size = 12), axis.text = element_text(size = 10))

# Scatter plot for PhysActiveDays vs. log(BMI)
plot_physactive_bmi <- ggplot(nhanes_data, aes(x = PhysActiveDays, y = log(BMI))) +
  geom_point() +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(title = "Figure20: Physical Active Days vs. log(BMI)", x = "PhysActiveDays", y = "log(BMI)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14), axis.title = element_text(size = 12), axis.text = element_text(size = 10))

# Combine all plots into a single figure using patchwork
library(patchwork)
combined_plot <- (plot_age_bmi | plot_gender_bmi) / 
                 (plot_poverty_bmi | plot_sleep_bmi) / 
                 plot_physactive_bmi +
  plot_annotation(
    title = "Diagnostic and Predictor Relationships in the Linear Model",
    caption = "Figure: Diagnostic and variable relationships from the linear model for log(BMI).",
    theme = theme(plot.caption = element_text(hjust = 0.5, size = 10))
  )

# Display the combined plot
combined_plot

@@

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

@@

# Additional data details

@@

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}

@@

```{r}
#| echo: false
#| label: fig-model-diagnostics
#| fig-cap: "Diagnostic and Predictor Relationships in the Linear Model for log(BMI)."
#| fig.width: 8
#| fig.height: 12

@@

# Create each plot
# 1. Fitted vs Residuals
plot_fitted_residuals <- ggplot(data.frame(fitted_values, residual_values), 
                                aes(x = fitted_values, y = residual_values)) +
  geom_point(color = "blue", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Fitted vs Residuals", x = "Fitted Values", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 2. Fitted vs Standardized Residuals
plot_fitted_standardized <- ggplot(data.frame(fitted_values, sresidual_values), 
                                   aes(x = fitted_values, y = sresidual_values)) +
  geom_point(color = "red", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Fitted vs Standardized Residuals", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 3. Standardized Residuals Histogram
plot_residuals_hist <- ggplot(data.frame(sresidual_values), aes(x = sresidual_values)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Standardized Residuals", x = "Standardized Residuals", y = "Frequency") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 4. Residuals vs Poverty
plot_residuals_poverty <- ggplot(data.frame(nhanes_data$Poverty, residual_values), 
                                 aes(x = nhanes_data$Poverty, y = residual_values)) +
  geom_point(color = "green", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Poverty vs Residuals", x = "Poverty Index", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 5. Residuals vs Age
plot_residuals_age <- ggplot(data.frame(nhanes_data$Age, residual_values), 
                             aes(x = nhanes_data$Age, y = residual_values)) +
  geom_point(color = "orange", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Age vs Residuals", x = "Age", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 6. Residuals vs Gender
plot_residuals_gender <- ggplot(data.frame(nhanes_data$Gender, residual_values), 
                                aes(x = nhanes_data$Gender, y = residual_values)) +
  geom_boxplot(fill = "gray") +
  labs(title = "Gender vs Residuals", x = "Gender", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 7. Fitted vs Log(BMI)
plot_fitted_logBMI <- ggplot(data.frame(fitted_values, log_BMI = scatter_data$log_BMI), 
                             aes(x = fitted_values, y = log_BMI)) +
  geom_point(color = "purple", size = 0.3) +
  geom_abline(intercept = 0, slope = 1, color = "blue", linetype = "dashed") +
  labs(title = "Fitted vs log(BMI)", x = "Fitted Values", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 8. Q-Q Plot
plot_qq <- ggplot(data = data.frame(std_resid = rstandard(lm_model)), aes(sample = std_resid)) +
  stat_qq(color = "darkred", size = 0.3) +
  stat_qq_line(color = "blue") +
  labs(title = "Normal Q-Q Plot", x = "Theoretical Quantiles", y = "Standardized Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# Combine all plots into one big plot
big_plot <- (
  (plot_fitted_residuals | plot_fitted_standardized) /
  (plot_residuals_hist | plot_residuals_poverty) /
  (plot_residuals_age | plot_residuals_gender) /
  (plot_fitted_logBMI | plot_qq)
)

# Display the combined plot
big_plot

>>>>

The dataset used in this analysis is derived from the US National Health and Nutrition Examination Survey (NHANES), version 2.1.0, published in July 2015. This version of the dataset includes survey results collected between October 2009 and December 2010. NHANES is a long-running study conducted by the US National Center for Health Statistics (NCHS) that has been gathering health and nutrition data since the early 1960s. Since 1999, approximately 5,000 individuals from various age groups have been interviewed annually in their homes and undergone health examinations at mobile examination centers (MEC). 
Two datasets in this package were considered for this analysis: NHANESraw, which is the original raw data, and NHANES, a resampled version of the NHANES data. NHANESraw contains the original survey data with 20,293 observations and additional variables describing the sample weighting scheme, while NHANES is a simplified version with 10,000 resampled observations to account for oversampling effects. NHANES is used in this analysis due to its ability to reduce the potential biases from the complex survey design in NHANESraw. Additional details about this choice are provided in Appendix A. The data used here was originally compiled by Michelle Dalrymple from Cashmere High School and Chris Wild from the University of Auckland for educational purposes.

For the current study, the data was cleaned to focus on variables pertinent to the analysis of BMI. Specifically, variables such as BMI, poverty index, physical activity days, sleep hours, gender, and age were retained. After cleaning the missing values in the dataset, 3,573 observations remained.

The dataset was prepared, cleaned, and analyzed using R [@citeR] with the following libraries: tidyverse [@tidyverse] and dplyr [@dplyr] for data manipulation, ggplot2 [@ggplot2] and patchwork [@patchwork] for visualizations, and broom [@broom] for model summaries. The arrow [@arrow] library was used for efficient data storage and retrieval, while knitr [@knitr] facilitated report generation. The NHANES package [@nhanes] provided access to the dataset, and styler [@styler] was employed to ensure well-structured R code. Additionally, testthat [@testthat] and pointblank [@pointblank] were utilized for data validation and testing.

@@

# Create the summary table using kable
summary_table %>%
  kable()
```

```{r}
#| warning: false
#| message: false
#| label: fig-bmi-logbmi
#| fig-cap: "Distributions of BMI and log(BMI) in the NHANES dataset."
#| echo: false
#| fig.width: 8
#| fig.height: 4

@@

# Figure 1: BMI and log(BMI)
figure_bmi <- (hist_bmi | hist_log_bmi)
figure_bmi

@@

```{r}
#| warning: false
#| message: false
#| label: fig-predictor-distributions
#| fig-cap: "Distributions of predictor variables in the NHANES dataset, including Poverty, Physical Activity Days, Age, Sleep Hours, and Gender."
#| echo: false
#| fig.width: 8
#| fig.height: 12

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-modelresults
#| tbl-cap: "Summary statistics for variables in the NHANES dataset."

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-predictor-bmi
#| fig-cap: "Scatter plots of predictors vs. log(BMI) with fitted regression lines."
#| fig-subcap: ["Age vs. log(BMI)","Gender vs. log(BMI)", "Poverty vs. log(BMI)", "SleepHrsNight vs. log(BMI)","Physical Active Days vs. log(BMI)"]
#| layout-ncol: 3
#| fig.width: 8
#| fig.height: 12

# Scatter plot for Age vs. log(BMI)
plot_age_bmi <- ggplot(nhanes_data, aes(x = Age, y = log(BMI))) +
  geom_point(size = 0.3) +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(x = "Age", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# Box plot for Gender vs. log(BMI)
plot_gender_bmi <- ggplot(nhanes_data, aes(x = Gender, y = log(BMI))) +
  geom_boxplot(fill = "lightblue", color = "black") +  # Box plot
  labs(x = "Gender", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# Scatter plot for Poverty vs. log(BMI)
plot_poverty_bmi <- ggplot(nhanes_data, aes(x = Poverty, y = log(BMI))) +
  geom_point(size = 0.3) +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(x = "Poverty", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# Scatter plot for SleepHrsNight vs. log(BMI)
plot_sleep_bmi <- ggplot(nhanes_data, aes(x = SleepHrsNight, y = log(BMI))) +
  geom_point(size = 0.3) +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(x = "SleepHrsNight", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# Scatter plot for PhysActiveDays vs. log(BMI)
plot_physactive_bmi <- ggplot(nhanes_data, aes(x = PhysActiveDays, y = log(BMI))) +
  geom_point(size = 0.3) +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(x = "PhysActiveDays", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

plot_age_bmi
plot_gender_bmi
plot_poverty_bmi
plot_sleep_bmi
plot_physactive_bmi

@@

## What is done in this paper? {#sec-first-point}

1.5 page This paper explores the relationship between Body Mass Index (BMI) and several key lifestyle and demographic factors, including poverty, physical activity, age, sleep duration, and gender. The analysis is performed using a linear regression model, which allows us to estimate how each of these predictors influences BMI, as measured by its log-transformed values. The dataset used for this study is the cleaned data from the US National Health and Nutrition Examination Survey (NHANES), which contains a broad range of health-related information from a representative sample of the U.S. population.

The model aims to quantify the impact of each predictor on BMI while controlling for the influence of other variables. The regression analysis reveals significant relationships between BMI and all the predictors considered, with some factors having a negative impact (e.g., poverty, physical activity) and others having a positive one (e.g., age, sleep hours, gender). This model sheds light on the ways in which various lifestyle choices and demographic factors may contribute to BMI variation across different segments of the population.

The model used in this study, a linear regression model, is a fundamental yet powerful tool that helps to identify the strength and direction of the relationships between predictors and BMI. The results provide a valuable starting point for understanding how changes in lifestyle and demographic factors may affect BMI. However, while the model is useful, it is not without its limitations. For instance, it assumes that the relationships between the predictors and BMI are linear, which may not always be the case. Additionally, the low R-squared value suggests that other important factors influencing BMI may not have been included in the model.

## What is something that we learn about the world?

1.5 page One of the key findings from this paper is that poverty is negatively correlated with BMI, and the relationship is statistically significant. The negative coefficient for the poverty variable suggests that individuals in poverty tend to have higher BMI, on average, than those with higher income levels. This result is consistent with a growing body of literature that links lower socioeconomic status to obesity and poor health outcomes. It may seem counterintuitive at first—since poverty is often associated with limited access to healthy foods and exercise—but the correlation can be explained by several factors.

In many low-income communities, the availability of affordable healthy food options is limited, leading individuals to rely on cheaper, calorie-dense foods that contribute to weight gain. Additionally, individuals in poverty may face multiple barriers to physical activity, including the lack of access to safe recreational spaces or the time and energy needed to engage in exercise, due to long working hours or other stressors. Furthermore, the high cost of healthcare and medical services in lower-income populations may prevent individuals from seeking preventative care and managing conditions like obesity. Thus, this finding underscores the importance of addressing socioeconomic disparities as part of efforts to combat obesity and improve public health.

Moreover, this finding highlights the need for policy interventions that promote access to healthy food and physical activity in low-income communities. It also suggests that any interventions aimed at reducing BMI must consider not just individual behaviors, but the broader socioeconomic context in which these behaviors occur. Policymakers should prioritize interventions that reduce the barriers to healthy living for individuals in poverty, such as improving access to nutritious food, increasing public health education, and providing community resources for physical activity.

## What is another thing that we learn about the world?

1.5 page Another key takeaway from this paper is that sleep duration has a significant negative correlation with BMI, suggesting that individuals who get more sleep tend to have lower BMI. This relationship is in line with previous research that has examined the link between sleep and obesity. The coefficient for sleep hours per night shows that for each additional hour of sleep, BMI decreases slightly, indicating that sleep plays a role in regulating body weight. This finding has important implications for public health, as sleep duration is a modifiable lifestyle factor that could be targeted in weight management programs.

The relationship between sleep and BMI can be explained through several mechanisms. Poor sleep has been shown to disrupt the body's metabolism, leading to an increase in appetite, particularly for high-calorie foods. Sleep deprivation also affects the hormones that regulate hunger and satiety, increasing the production of ghrelin (the hunger hormone) while decreasing the production of leptin (the satiety hormone). These changes can lead to overeating and, consequently, weight gain. Furthermore, insufficient sleep can lead to fatigue, reducing the likelihood of physical activity and further contributing to weight gain. Therefore, improving sleep hygiene and encouraging sufficient sleep may serve as an effective strategy for reducing BMI and addressing obesity.

This result also emphasizes the importance of taking a holistic approach to health. In addition to focusing on diet and physical activity, interventions aimed at improving sleep quality could be a key component in the prevention and treatment of obesity. Public health campaigns that educate individuals about the importance of sleep, along with strategies to improve sleep hygiene (e.g., creating a conducive sleep environment, avoiding caffeine, establishing regular sleep routines), could complement existing efforts to promote healthy eating and physical activity.

@@

# Additional data details {#sec-data-details}

## Data Justication
The NHANES_raw dataset provides a comprehensive collection of data from the American National Health and Nutrition Examination Surveys (NHANES), including 20,293 observations spanning 75 variables. This dataset includes additional variables that describe the sample weighting scheme employed to account for the complex survey design. However, despite its breadth and richness, NHANES_raw introduces complexities in analysis due to the oversampling of certain subpopulations, such as racial minorities. This sampling approach necessitates the use of survey weights and other design-specific parameters to obtain unbiased estimates and valid statistical inferences.

Naïve analysis of NHANES_raw—without accounting for the sample design and weights—can lead to significant biases and erroneous conclusions. For instance, the racial composition in NHANES_raw does not represent the true demographic distribution of the U.S. population due to oversampling adjustments. As a result, researchers must account for these weights in any analysis to avoid overstating significance levels and generating inaccurate results. This makes NHANES_raw unsuitable for straightforward analysis in contexts where the primary goal is to demonstrate simple statistical relationships, especially for educational purposes.

In contrast, the NHANES dataset, which is a resampled subset of NHANES_raw, addresses the complexities associated with the oversampling design. With 10,000 observations and the same 75 variables, NHANES is simplified to approximate a simple random sample from the U.S. population. This resampling process removes the need for researchers to manually account for the oversampling effects, enabling analyses that are easier to interpret and less prone to methodological errors for educational purposes.

The decision to use NHANES in this paper was driven by its appropriateness for demonstrating relationships between body mass index (BMI) and lifestyle, poverty, age, and gender. NHANES allows for clear and reliable analyses without requiring specialized expertise in survey design or sampling weights, making it ideal for this study’s educational focus. The dataset enables meaningful exploration of public health questions while minimizing the risk of misinterpretation due to unaddressed biases in the sampling framework.

It is important to acknowledge that NHANES is adapted for educational purposes and is not intended for formal research. Researchers conducting advanced studies should rely on the original NHANES_raw data and adhere to the prescribed sampling and weighting procedures to ensure the validity of their findings. Future research could extend the current analysis by using NHANES_raw in combination with survey weighting techniques to explore population-level health trends and disparities with greater accuracy.

# Surveys, Sampling, and Observational Data {#sec-sampling-details}

@@

@@

```{r}
#| echo: false
#| label: fig-model-diagnostics
#| fig-cap: "Diagnostic and Predictor Relationships in the Linear Model for log(BMI)."
#| fig-subcap: ["Fitted vs Residuals", "Fitted vs Standardized Residuals", "Histogram of Standardized Residuals", "Poverty vs Residuals", "Age vs Residuals", "Gender vs Residuals", "Fitted vs log(BMI)", "Normal Q-Q Plot"]
#| layout-ncol: 4
#| fig.width: 12
#| fig.height: 18

@@

# Create each plot
# 1. Fitted vs Residuals
plot_fitted_residuals <- ggplot(data.frame(fitted_values, residual_values), 
                                aes(x = fitted_values, y = residual_values)) +
  geom_point(color = "blue", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Fitted Values", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 2. Fitted vs Standardized Residuals
plot_fitted_standardized <- ggplot(data.frame(fitted_values, sresidual_values), 
                                   aes(x = fitted_values, y = sresidual_values)) +
  geom_point(color = "red", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 3. Standardized Residuals Histogram
plot_residuals_hist <- ggplot(data.frame(sresidual_values), aes(x = sresidual_values)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(x = "Standardized Residuals", y = "Frequency") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 4. Residuals vs Poverty
plot_residuals_poverty <- ggplot(data.frame(nhanes_data$Poverty, residual_values), 
                                 aes(x = nhanes_data$Poverty, y = residual_values)) +
  geom_point(color = "green", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Poverty Index", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 5. Residuals vs Age
plot_residuals_age <- ggplot(data.frame(nhanes_data$Age, residual_values), 
                             aes(x = nhanes_data$Age, y = residual_values)) +
  geom_point(color = "orange", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Age", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 6. Residuals vs Gender
plot_residuals_gender <- ggplot(data.frame(nhanes_data$Gender, residual_values), 
                                aes(x = nhanes_data$Gender, y = residual_values)) +
  geom_boxplot(fill = "gray") +
  labs(x = "Gender", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 7. Fitted vs Log(BMI)
plot_fitted_logBMI <- ggplot(data.frame(fitted_values, log_BMI = scatter_data$log_BMI), 
                             aes(x = fitted_values, y = log_BMI)) +
  geom_point(color = "purple", size = 0.3) +
  geom_abline(intercept = 0, slope = 1, color = "blue", linetype = "dashed") +
  labs(x = "Fitted Values", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 8. Q-Q Plot
plot_qq <- ggplot(data = data.frame(std_resid = rstandard(lm_model)), aes(sample = std_resid)) +
  stat_qq(color = "darkred", size = 0.3) +
  stat_qq_line(color = "blue") +
  labs(x = "Theoretical Quantiles", y = "Standardized Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

plot_fitted_residuals
plot_fitted_standardized
plot_residuals_hist
plot_residuals_poverty
plot_residuals_age
plot_residuals_gender
plot_fitted_logBMI
plot_qq
